# Off scope - Analyzing FACS data in VR

Analyzing FACS data in VR is not the original scope of CellexalVR.
In line with that we need some extra packages installed to be able to run this.

```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
if (!require("flowCore", quietly = TRUE))
	BiocManager::install("flowCore")
if (!require("umap", quietly = TRUE))
	install.packages( c( 'umap' ) )

```

What we will do here is load FACS data from a fsc file, calculate a umap from the data and load this data into a cellexalvrR object.
We will 'load' this data by replacing the existing data in the example cellexalObj.


First we will load the data and claculate a 3D umap.

```{r}
suppressWarnings(suppressMessages(library(flowCore)))
## let's use the flowCore example file
file.name <- system.file("extdata","0877408774.B08",package="flowCore")
x <- read.FCS( file.name, transformation=FALSE)

pca = prcomp( t(x@exprs) )

suppressWarnings(suppressMessages(library(umap)))
umap =  umap(pca$rotation, n_components=3)
```

Next we will populate the cellexalObj with the FACS data and the 3D umap.

```{r}
suppressWarnings(suppressMessages(library(cellexalvrR)))

cellexalObj@index = apply( x@exprs, 2,log)

if ( length( which( is.na(cellexalObj@index))) > 0 ) {
	    cellexalObj@index[ which( is.na(cellexalObj@index)) ] = 0
}
if ( length(which(! is.finite(cellexalObj@index))) > 0 ) {
	cellexalObj@index[ which(! is.finite(cellexalObj@index)) ] = 0
}

cellexalObj@data = Matrix( t ( cellexalObj@index ), sparse=T)
colnames(cellexalObj@data) = paste('cell', 1:ncol(cellexalObj@data) ,sep ="_" )
rownames(cellexalObj@data) = stringr::str_replace_all(as.vector(rownames(cellexalObj@data)) ,"\\s\\s*", "_" )

rownames(cellexalObj@index) = colnames(cellexalObj@data)
colnames(cellexalObj@index) = rownames(cellexalObj@data)

cellexalObj@drc = list('umap' = umap$layout )
rownames(cellexalObj@drc$umap)=colnames(cellexalObj@data)

cellexalObj@meta.cell = cbind( "FACS_data" = rep(1, ncol(cellexalObj@data)), "saveKeep" = rep(1, ncol(cellexalObj@data)) )
rownames(cellexalObj@meta.cell)=colnames(cellexalObj@data)

cellexalObj = check(cellexalObj)

```

This is what we got here:

```{r}
cellexalObj
```


And finally we can export the data for the VR process:

```{r}
export2cellexalvr( cellexalObj ,"FACS_Data")

list.files(opath)

```

```{r, include=FALSE, eval=T}
rm(cellexalObj)
```

The folder FACS_data can be copied to the CellexalVR data folder and is available in the next VR session.

